### 2016-05-31 main.tex
first commit.

% Required packages
%\RequirePackage{fix-cm}

% Document Settings
\documentclass[a4paper,review]{elsarticle}

% Packages
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[margin=3cm]{geometry}

% Journal Name
\journal{Physica A}

% START: The Document
\begin{document}

\begin{frontmatter}

% Long Title
\title{Visibility Graph Analysis of the 2005-2015 Earthquake Sequence in Northern Iran}

% Authors
%\author[label1]{Naeem Khoshnevis\corref{cor1}}
%\author[label1,label2]{Shima AzizzadehRoodpish}
%\cortext[cor1]{Corresponding author. Tel: \\
 %                    Email address: nkhshnvs@memphis.edu (N.Khoshnevis)}

% Affliations
%\address[label1]{Center for Earthquake Research and Information}
%\address[label2]{Department of Civil Engineering}

\input{abstract}

\begin{keyword}
Visibility Graph \sep Earthquake Magnitude \sep North Iran \sep b-value

\end{keyword}

\end{frontmatter}

%% \linenumbers

%% Article Contents

\input{content}
\bibliographystyle{elsarticle-num-names}
\bibliography{VG}

% Bibliography

%\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

%\bibitem{}

%\end{thebibliography}

% End document
\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.