### 2016-07-30 acknowledgements.tex
Adding the synthetic data of Telesca 2014 (POne) paper. And general review.

\section{Acknowledgements}
\noindent
We greatly appreciate the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for maintaining and providing access to the seismic catalog. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunsent Laws (SB 1510 and HB 1608, 2015{2016).