### 2016-08-11 title-and-authors.tex
formating for srl

\pretitle{
    \vspace{-30pt}
    \begin{flushleft}
    \usefont{T1}{phv}{bc}{it}
    \Huge
}

\posttitle{
    \end{flushleft}
}

\preauthor{
    \begin{flushleft}
    \vspace{3pt}
    \usefont{T1}{phv}{bc}{n}
    \LARGE
}

\postauthor{
    \end{flushleft}
    \vspace{0.3em}
    \section{Abstract}
    \vspace{-3em}
}

\predate{
    \begin{flushleft}
    \makeatletter
}

\postdate{
    \end{flushleft}
}

\title{%
    Analysis of 2005--2015 Earthquake Sequence in Northern Iran Using Visibility Graph Approach
}

\makeatletter
\if@twocolumn
    \date{}
    \renewcommand*\@fnsymbol[1]{\the#1}
    \author{by Naeem Khoshnevis, Shima Azizzadeh-Roodpish, Ricardo Taborda, and Luciano Telesca}
\else
    \author{
    	Naeem Khoshnevis,\myaffil{1,$\ast$} Shima Azizzadeh-Roodpish,\myaffil{1,2} Ricardo Taborda\myaffil{1,2} and Luciano Telesca\myaffil{3}\\
    	\normalsize\normalfont    	    	
        \vspace{20pt}
    	1.~
    	Center for Earthquake Research and Information\\
    	\hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        2.~
        Department of Civil Engineering\\
        \hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        3.~
        Institute of Methodologies for Environmental Analysis\\
        \hspace{1.1em} National Research Council, Tito, Italy\\
        \vspace{20pt}
        *.~
    	Corresponding author\\
    	\hspace{1.1em} Email: nkhshnvs@memphis.edu\\
        \vspace{40pt}
    }
\fi
\makeatother