### 2016-08-19 acknowledgements.tex
started with results section, added improved figure 6

\section{Acknowledgments}

This research was possible thanks to support from the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunset Laws (SB 1510 and HB 1608, 2015--2016).