### 2016-08-19 data-and-resources.tex

\section{Data and Resources}

Data used to compile the seismic catalog employed in the analysis was obtained using the Web browser search application from the International Institute of Earthquake Engineering and Seismology, IIEES (\url{http://www.iiees.ac.ir/en/eqcatalog/}, last accessed September 2016). The raw versions of figures were prepared using MATLAB\textsuperscript{\textregistered}, release 13b; and the Generic Mapping Tools, version 5.1 \citep[\url{http://gmt.soest.hawaii.edu;}][]{Wessel_2013_EOS}. Final versions of figures were prepared using Adobe\textsuperscript{\textregistered} Illustrator\textsuperscript{\textregistered}.