### 2016-08-21 title-and-authors.tex

\title{%
    Analysis of the 2005--2015 Earthquake Sequence in Northern Iran Using the Visibility Graph Method
}

\makeatletter
\if@twocolumn
    \date{}
    \renewcommand*\@fnsymbol[1]{\the#1}
    \author{by Naeem Khoshnevis, Ricardo Taborda, Shima Azizzadeh-Roodpish, and Luciano Telesca}
\else
    \author{
        Naeem Khoshnevis,\myaffil{1,$\ast$} Ricardo Taborda\myaffil{1,2} Shima Azizzadeh-Roodpish,\myaffil{1,2} and Luciano Telesca\myaffil{3}\\
        \normalsize\normalfont    	    	
        \vspace{20pt}
        1.~
        Center for Earthquake Research and Information\\
        \hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        2.~
        Department of Civil Engineering\\
        \hspace{1.1em} The University of Memphis, Memphis, TN 38152, U.S.A.\\
        \vspace{20pt}
        3.~
        Institute of Methodologies for Environmental Analysis\\
        \hspace{1.1em} National Research Council, Tito, Italy\\
        \vspace{20pt}
        *.~
        Corresponding author\\
        \hspace{1.1em} Email: nkhshnvs@memphis.edu\\
        \vspace{40pt}
    }
\fi
\makeatother