### 2016-08-22 conclusion.tex
done with conclusions and final fixes

\section{Conclusions}

We studied the seismicity of the three main seismic regions in northern Iran in the time period between January 2005 and December 2015 using an approach based on the visibility graph method. We tested the applicability of this method for the specific region of interest and in reference to previous results from similar studies. The results confirm previous observations about the correlation that exists between the connectivity parameter $k$, the magnitude of the events in the sequence, the slope of this relationship or $k$-$M$ slope, and the $b$-value from the Gutenberg-Richter law. We obtained mathematical expressions for the region of interest as well as for so-called universal data collected from this study and previous additional analysis for other regions as well as data from experiments and found the relationships to have a good level of similarity. This is indicative of the general nature of the relationship between the $k$-$M$ slope and the $b$ value. We also explored the potential relationships that can be drawn from a time-dependent visibility graph analysis and the variation of the seismicity in the region of interest. We found there may be a potential relationship between the the $b$ value and the occurrence of earthquakes as well as with the graph's mean interval connectivity time parameter <$T_c$>, but additional research in regions with stronger events may be necessary before drawing stronger conclusions in this regard. The method used here, nonetheless, seems to provide an alternative and interesting approach to the analysis of the seismicity of a region.

% We studied the earthquake sequence of three tectonic seismic regions of northern Iran. Results manifest the strong relationship between  $k-M$  slope and Gutenberg-Richter parameters. The variations of  $k-M$  slope and  b-value  with time in sliding windows have similar behavior and decline before large earthquakes. A significant reduction of  $<T_c>$  value before large earthquakes was observed. The higher coefficient of variation of  $k-M$  slope in comparison with  b-value  suggests that the  $k-M$  slope is a better parameter for studying the magnitude time series due to higher sensitivity to window size and threshold magnitude than  is found in the $b-value$. Sensitivity analysis revealed that the VG parameters behave independently than threshold magnitude and number of events. In general,  the $k-M$  and  $b-value$  relationship preserves the behavior with the variation in the window size and threshold magnitude, and could be an alternative approach for clustering earthquake sequences based on different tectonic seismic zones.