### 2017-06-20 acknowledgements.tex
working revised version for paag

\section{Acknowledgments}
\label{sec:acknowledgments}

\myrevision{We are thankful for the comments and suggestions of two anonymous reviewers, which helped us improve the original manuscript. The data used to compile the seismic catalog employed in this analysis was obtained from the International Institute of Earthquake Engineering and Seismology, IIEES (\url{http://www.iiees.ac.ir/en/eqcatalog/})}. This research was possible thanks to support from the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunset Laws (SB 1510 and HB 1608, 2015--2016).